A Quasi-Genuine FIFO Total Order Multicast Primitive

### 2011-06-10 trunk/textos/optpaxos/draftv3.tex
paxos 3 delta image

We need to mention the clock deviation in this assumption because, even if a process waits for a period longer than the time needed for a message $m$ to arrive since it was sent, it might be the case that a clock deviation between the processes caused the timestamp of $m$ to be too low and lead to a violation of the timestamp order.

@@

In the context of our \cms{} primitive, the first communication step of Paxos -- forwarding a message $m$ to the leader -- is already done by means of \rmc{}ing each message to a set of processes that includes every process in $m$'s group of origin. The last phase, which consists of every learner receiving the confirmation of a majority of acceptors, can also include the $\delta$ included in the conservative message delivery time we analysed before: instead of waiting for a message to be decided and only then notify other groups, the acceptors of a group can also send the confirmation to processes in other groups, so that also they can infer the first group's decision right away. This implies that, for each consensus instance run within a group $G$, each the processes in the groups of $receiversFrom(G)$ would be a learner for that instance. Figure \ref{fig:paxosmegalearners} illustrates this: the group $G$ is delivering a message $m$ and the last phase of the consensus protocol already notifies the processes in $G' \in receiversFrom(G)$.

>>>>

We need to mention the clock deviation in this assumption because, even if a process waits for a period longer than the time needed for a message $m$ to arrive since it was sent, it might be the case that a clock difference between the processes caused the timestamp of $m$ to be too high and delivering it might lead to a violation of the timestamp order once some other message, with a lower timestamp, arrives. More formally, if $p \in G$, we can define $w(p)$ as $max_{p' \in sendersTo(G)}(\delta(p,p')+\epsilon(p,p'))$, where $\delta(p,p')$ is the maximum time a message takes to go all the way from $p$ to $p'$ and $\epsilon(p,p')$ is the difference between the clocks of $p$ and $p'$, so that clock deviations can be accounted for\footnote{If this difference is less than zero, it means that the clock value in $p'$ is higher than that in $p$, so $p$ actually has to wait less time for messages from $p'$, as they will have higher timestamps because of the clock deviation.}.

@@

In the context of our \cms{} primitive, the first communication step of Paxos -- forwarding a message $m$ to the leader -- is already done by means of \rmc{}ing each message to a set of processes that includes every process in $m$'s group of origin. The last phase, which consists of every learner receiving the confirmation of a majority of acceptors, can also include the $\delta$ included in the conservative message delivery time we analysed before: instead of waiting for a message to be decided and only then notify other groups, the acceptors of a group can also send the confirmation to processes in other groups, so that also they can infer the first group's decision right away. This implies that, for each consensus instance run within a group $G$, each the processes in the groups of $receiversFrom(G)$ would be a learner for that instance. Figure \ref{fig:paxosmanylearnergroups} illustrates this: process $p_1$ from group $G$, whose leader is $p_3$, \cms{}s $m$; in the last phase of the consensus protocol, the processes in $G' \in receiversFrom(G)$ receive the confirmation message from the acceptors of $G$, so no extra communication step is needed to notify them about the final timestamp of $m$.

As $m$ was \rmc{}ed in the beginning, there is no need to send it again in the consensus instance. Instead, only the timestamp of $m$ is proposed, since this is the information that must be agreed upon -- and which may change if the optimistic assumption does not hold and some other message of higher timestamp is delivered before $m$. In any case, as all messages agreed upon in $G$ are notified to $G'$, also the processes of $G'$ are able to infer the final timestamp of each message from $G$.

Also, as we are making an optimistic assumption -- that each $p$ waits for a time $w(p)$ long enough so that no timestamps must be changed -- for the message delivery, the barriers are being requested to other groups in parallel, since $m$ is also \rmc{} to $blockers(m)$ in the first communication step of our protocol. If such assumption holds, the conservative delivery of the messages will take place $w(p) + 2\delta$ after it has been sent. If the clocks are perfectly synchronized and $w(p)$ does correspond to the maximum interprocess communication delay $\delta$, the time needed to conservatively deliver a message using this algorithm would be $3\delta$.

\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{images/paxos3d}
  \caption{}
  \label{fig:paxosmanylearnergroups}
\end{figure}
